Data Encryption Standard Implementation 4003-482-01 RIT

### 2010-05-06 paper/paper.tex
Add more stuff

\usepackage{graphicx}
\usepackage{multicol}
% \usepackage{acronym}
\usepackage[bookmarks=false,colorlinks=false,linkcolor={black},pdfstartview={XYZ null null 1.22}]{hyperref}
\usepackage{url}

@@

\begin{abstract}
  {DES} is the 56-bit block cipher that was once used to encrypt all US Government non-classified documents.
\end{abstract}

@@

\part{The Cryptographic Primitive}

\section{Examples}
Each was encrypted using the key $24681012$.
\subsection{Example 1}
\textit{Hello}
\\\indent Original:
\begin{verbatim}
   48 65 6C 6C 6F 0A
\end{verbatim}
\indent Encrypted:
\begin{verbatim}
   94 93 81 0E 62 47 30 86
\end{verbatim}

@@

@@

The last function, f(), also known as the Feistel function, is the heart of the encryption.  It first uses mixer() to do the expansion permutation, then xor's the result with the sub key. This output is split into 8-bit block and each block is sent through an S-Box. The output from the S-Boxes is then concatenated run through mixer() using the primitive permutation.
\\ \\ \\
We chose to implement DES completely in $C++$ for this project.
Both of the members our group were knowledgeable and comfortable
with $C++$ and in the $C$ world in general. We felt we would have a more
generally usable and useful implementation if we went with $C++$ over
another language (Java, Python, ... ). We attempted to take advantage of ac{OOP} but
were careful not to over do it and dive into a downword spiral of subclasses and
analysis of class cohesion. Our $DES$ implementation is simply a class, with actions of encrypting and
decrypting. All state information is organized within the class. We then
implemented a $BlockMode$ class which implements a block stream cipher, which is
just a wrapper around $DES$. The user is then provided with
two small utility functions $encrypt$ \and $decrypt$.

@@

   "We should forget about small efficiencies, say about 97\% of the time: premature optimization is the root of all evil"

@@

.... test output ...
\end{verbatim}
}

@@

In order to encrypt a file using out implementation the user must run the executable \textbf{encrypt\_opt}.
\begin{verbatim}

@@

The user has a similar choice for decrypting, both an optimized and original implementation exist.
\begin{verbatim}

@@

  \bibitem{ gtest }{ \href{ http://code.google.com/p/googletest/ }{ Google Unit Testing Framework } }
  \bibitem{ spec  }{ \href{ http://csrc.nist.gov/publications/fips/fips46-3/fips46-3.pdf }{ Data Encryption Standard Specification } }
  \bibitem{ wiki  }{ \href{ http://en.wikipedia.org/wiki/Data_Encryption_Standard }{ Wikipedia: Data Encryption Standard } }
  \bibitem{ block }{ \href{ http://en.wikipedia.org/wiki/Block_cipher_modes_of_operation }{ Wikipedia: Block Cipher Modes } }
  \bibitem{ desge }{ \href{ http://www.comms.scitech.sussex.ac.uk/fft/crypto/des_algorithm_details.txt}{ How to implement the Data Encryption Standard (DES) } }
  \bibitem{ desg2 }{ \href{ http://orlingrabbe.com/des.htm }{ DES Illustrated } }
  \bibitem{ gprof }{ \href{ http://linux.die.net/man/1/gprof }{ GProf Manual Page } }
  \bibitem{ g++ }{ \href{ http://linux.die.net/man/1/g++ }{ GCC Manual Page } }

>>>>

\usepackage{graphicx}
\usepackage{color}
\usepackage{multicol}
% \usepackage{acronym}
\usepackage[bookmarks=false,colorlinks=false,linkcolor={black},pdfstartview={XYZ null null 1.22}]{hyperref}
\usepackage{url}

@@

%\begin{abstract}
%  {DES} is the 56-bit block cipher that was once used to encrypt all US Government non-classified documents.
%$\end{abstract}

@@

\columnseprule .4pt

\definecolor{light-gray}{gray}{0.95}

\def \columnseprulecolor{ \color{light-gray} }

\section{The Cryptographic Primitive}

  {DES} is the 64-bit block cipher that was once used to encrypt all US Government non-classified documents.
  It was developed as a collaboration between NSA and IBM in the 1970's and was eventually selected
  by the National Institute of Standards and Technologies and was  standardised on January 15, 1977 as FIPS-46.
  As mentioned before the cipher has a 64-bit data block and uses a 64-bit key block, it should be noted though that
  only 56 bits of the key are ever actually used. The algorithm is based around sixteen rounds each with it's own
  shortened subkey (48 bits) based on the original key block. At the beginning of the cipher the data bock is permuted
  and is then inversely permuted

\section{Examples}
Each was encrypted using the key $24681012$.

\begin{multicols}{3}

\hbadness=10000

  \subsection{Example 1}
  \textit{Hello}
  \\\indent Original:
  \begin{verbatim}
     48 65 6C 6C 6F 0A
  \end{verbatim}
  \indent Encrypted:
  \begin{verbatim}
     94 93 81 0E 62 47 30 86
  \end{verbatim}

@@

\end{multicols}

@@

The last function, f(), also known as the Feistel function, is the heart of the encryption.  It first uses mixer() to do the expansion permutation, then xor's the result with the sub key. This output is split into 8-bit block and each block is sent through an S-Box. The output from the S-Boxes is then concatenated run through mixer() using the primitive permutation.

@@

\begin{quote}
   "We should forget about small efficiencies, say about 97\% of the time: premature optimization is the root of all evil"
\end{quote}

@@

... test output ...
\end{verbatim}
}

@@

\noindent In order to encrypt a file using our implementation the user must run the executable \textbf{encrypt\_opt}.
\begin{verbatim}

@@

\noindent The user has a similar choice for decrypting, both an optimized and original implementation exist.
\begin{verbatim}

@@

  \bibitem{gtest}{ \href{ http://code.google.com/p/googletest/ }{ Google Unit Testing Framework } }
  \bibitem{ spec  }{ \href{ http://csrc.nist.gov/publications/fips/fips46-3/fips46-3.pdf }{ Data Encryption Standard Specification } }
  \bibitem{ wiki  }{ \href{ http://en.wikipedia.org/wiki/Data_Encryption_Standard }{ Wikipedia: Data Encryption Standard } }
  \bibitem{ block }{ \href{ http://en.wikipedia.org/wiki/Block_cipher_modes_of_operation }{ Wikipedia: Block Cipher Modes } }
  \bibitem{ desge }{ \href{ http://www.comms.scitech.sussex.ac.uk/fft/crypto/des_algorithm_details.txt}{ How to implement the Data Encryption Standard (DES) } }
  \bibitem{ desg2 }{ \href{ http://orlingrabbe.com/des.htm }{ DES Illustrated } }
  \bibitem{ gprof }{ \href{ http://linux.die.net/man/1/gprof }{ GProf Manual Page } }
  \bibitem{ g++ }{ \href{ http://linux.die.net/man/1/g++ }{ GCC Manual Page } }
